---
title: "GitHub Petitions Regarding Removal of RMS: Contributor Comparison"
author: "Wm Salt Hale  \n<www.altsalt.net>"
date: "March 26, 2021"
lang: en-US
fontsize: 12pt
# documentclass: article
# classoption: oneside
# geometry: margin=1in
# linkcolor: black
# urlcolor: blue
# citecolor: black
output:
  bookdown::pdf_document2:
    toc: false
#    toc: true
#    toc_depth: 2
#    number_sections: true
#    fig_caption: yes
    fig_width: 6
    fig_height: 4
    latex_engine: xelatex
    keep_tex: false
#    keep_tex: true
    citation_package: biblatex
    includes:
      in_header: preamble.tex
#bibliography: resources/refs.bib
biblio-style: apa
biblatexoptions: [backend=biber,natbib=true]
link-citations: yes
editor_options:
  chunk_output_type: console
---

```{r Global, include=FALSE, echo=FALSE, message=FALSE, warnings=FALSE}
# Clean Workspace
rm(list = ls())

# Global Options
knitr::opts_chunk$set(echo=FALSE, error=FALSE, message=FALSE, warning=FALSE)

# Global Variables
gen_plots_toggle <- FALSE

```

```{r Libraries, include=FALSE}
# Load Libraries
required_pkgs <- c("here", "RColorBrewer", "tidyverse", "scales",
                   "knitr", "bookdown", "kableExtra")
installed_pkgs <- installed.packages()
sapply(required_pkgs, function(p) {
  if(!p %in% installed_pkgs[,1]) {
    install.packages(p)
  }
})
# https://stackoverflow.com/questions/8175912/load-multiple-packages-at-once/8176099#8176099
lapply(required_pkgs, require, character.only = TRUE)

here::i_am('rms-letter-comparison.Rmd')
here()

```

```{r Helper Functions, include=FALSE}
# Helper Functions
source(here('code', '99_helpers.R'))

```

```{r Load Data, include=FALSE}
# Load Data
df <- readRDS(here('data', 'dataset.Rds'))
df_facts <- readRDS(here('data', 'facts.Rds'))
df_tables <- readRDS(here('data', 'tables.Rds'))

```

On March 22, 2021 the former president of the Free Software Foundation (FSF) Richard M. Stallman (rms) announced that he would be returning to the FSF board.\footnote{\url{https://twitter.com/nixcraft/status/1373905399707955202}} This comes 18 months after he stepped down, due to public pressure regarding, among other things, statements relating to the involvement of Marvin Minsky and Jeffrey Epstein.\footnote{\url{https://itsfoss.com/richard-stallman-controversy}}

On March 23, 2021 a letter was published online calling for the removal of rms from all leadership positions within the Free/Libre/Open Source Software community.\footnote{\url{https://rms-open-letter.github.io/}} Shortly thereafter, a counter letter, in support of the return of rms was published.\footnote{\url{https://rms-support-letter.github.io/}}

Each of these letters called for readers to add their signatures via a pull request (PR) using GitHub, one of the largest hosts for FLOSS projects. As of `r format(Sys.time(), "%T %Z")` on `r format.date(Sys.time())`, the letter requesting removal has received `r f(df_facts$orig_rr_signers)` signatures, while the letter in support of remaining has received `r f(df_facts$orig_sr_signers)`.

Below is a cursory analysis comparing contributors to each letter. The data used was acquired from GitHub via their official API, accessed using the PyGithub library.\footnote{\url{https://github.com/PyGithub/PyGithub}} It was then processed in R, making use of tools from the tidyverse library.\footnote{\url{https://www.tidyverse.org/}}

The documentation provided for this project is released under a Creative Commons Attribution-ShareAlike 4.0 International License (CC BY-SA 4.0) https://creativecommons.org/licenses/by-sa/4.0/. The code provided for this project is released under the GNU General Public License version 3 (GNU GPLv3) https://www.gnu.org/licenses/gpl-3.0. The data was collected from GitHub. All data provided for this project is subject to the terms outlined on this page: https://docs.github.com/en/github/site-policy/github-terms-of-service


```{r breakdowntable, results='asis'}
printable_table <- round(df_tables$breakdown_table, digits=2) %>%
  kbl(booktabs=TRUE, escape=TRUE, caption='User breakdown of contributors on GitHub to letters calling for the removal or support of, Richard M. Stallman, respectively.') %>%
  kable_styling(latex_options=c('striped'))

print(printable_table)
```

<!--
```{r, eval=FALSE}
# Format facts table for plotting
#facts_table_orig <- facts_table
facts_table$Total <- NULL
facts_table <- rownames_to_column(facts_table, 'metric')
facts_table$metric <- factor(facts_table$metric, levels=facts_table$metric)

facts_table <- facts_table %>%
  pivot_longer(!'metric', names_to='letter', values_to='observation')

# Plot facts table
gg <- ggplot(facts_table, aes(x=metric, y=observation)) +
  geom_bar(aes(fill=letter), stat='identity', position='fill') +
  geom_hline(yintercept=0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=-40, hjust=.1)) +
  labs(fill='Letter to:',
       x=element_blank(),
       y=element_blank()) +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(facts_table$metric))) +
  scale_y_continuous(labels = percent)
```

```{r factsbarplot, results='asis', eval=FALSE}
print(gg)
```
-->

\appendix

# Organizational messages regarding the current events:
- [CommitChange](https://twitter.com/wwahammy/status/1374771022289854465)
- [Electronic Frontier Foundation (EFF)](https://www.eff.org/deeplinks/2021/03/statement-re-election-richard-stallman-fsf-board)
- [Free Software Foundation Europe (FSFE)](https://fsfe.org/news/2021/news-20210324-01.html)
- [Mozilla](https://twitter.com/mozilla/status/1374513444838199304)
- [Open Source Initiative (OSI)](https://opensource.org/OSI_Response)
- [Outreachy](https://www.outreachy.org/blog/2021-03-23/fsf-participation-barred/)
- [Red Hat](https://www.redhat.com/en/blog/red-hat-statement-about-richard-stallmans-return-free-software-foundation-board)
- [Software Freedom Conservancy](https://sfconservancy.org/blog/2021/mar/23/outreachy-fsf/)
- [The Tor Project](https://twitter.com/torproject/status/1374754834050654212)


```{r}
#saveRDS(df, file=here('data', 'dataset.Rds'))
#saveRDS(df_facts, file=here('data', 'facts.Rds'))
#saveRDS(df_tables, file=here('data', 'tables.Rds'))
```
